### 2018-05-16 paper.tex
include experiments part

\documentclass[12pt]{article}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{tcolorbox}
\tcbuselibrary{skins}
\usepackage{lipsum}

\usepackage{listings}

%%
%% Julia definition (c) 2014 Jubobs
%%
\lstdefinelanguage{Julia}%
  {morekeywords={abstract,break,case,catch,const,continue,do,else,elseif,%
      end,export,false,for,function,immutable,import,importall,if,in,%
      macro,module,otherwise,quote,return,switch,true,try,type,typealias,%
      using,while},%
   sensitive=true,%
   alsoother={\$},%
   morecomment=[l]\#,%
   morecomment=[n]{\#=}{=\#},%
   morestring=[s]{"}{"},%
   morestring=[m]{'}{'},%
}[keywords,comments,strings]%

\lstset{%
    language         = Julia,
    basicstyle       = \ttfamily,
    keywordstyle     = \bfseries\color{blue},
    stringstyle      = \color{magenta},
    commentstyle     = \color{ForestGreen},
    showstringspaces = false,
}

\usepackage[linesnumbered]{algorithm2e}
\makeatletter
\renewcommand{\@algocf@capt@plain}{above}
\makeatother
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{arrows}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\let\emptyset\varnothing
\definecolor{hotpink}{rgb}{0.9,0,0.5}
\definecolor{deepgray}{gray}{0.35}
\definecolor{deepgray2}{gray}{0.25}
\definecolor{deepgray3}{gray}{0.10}
\definecolor{lightgray}{gray}{0.95}
\definecolor{lightgray2}{gray}{0.85}
\definecolor{purple1}{RGB}{239,229,244}
\definecolor{purple2}{RGB}{216,191,216}
\definecolor{lightblue}{rgb}{0.73,0.33,0.83}
\definecolor{lightpurple}{rgb}{.8,.2,.8}
\definecolor{textcolor}{rgb}{0,0,5}
\definecolor{blue1}{RGB}{187,217,238}
\definecolor{blue2}{RGB}{235,244,250}
\definecolor{yellow1}{RGB}{255,255,102}
\definecolor{blue3}{RGB}{63,40,96}
\definecolor{red1}{RGB}{255, 102, 102}
\definecolor{green1}{RGB}{102, 255, 102}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\theoremstyle{definition}
\def\proof{\par{\bf Proof}. \ignorespaces}
\def\qedsymbol{\vbox{\hrule\hbox{%
                     \vrule height1.3ex\hskip0.8ex\vrule}\hrule}}
\def\endproof{\qquad\qedsymbol\medskip\par}

% \usepackage{booktabs} % For formal tables
% \usepackage{cite}
\usepackage{url}
% \newtheorem{definition}{Definition}[section]
% \newtheorem{theorem}{Theorem}[section]
% \usepackage{algorithmic}
% \usepackage[cmex10]{amsmath}
\usepackage{amssymb}

\title{A Closer Look at Network Flow on Evolving Graphs}

\author{Jiahao Chen \thanks{
Capital One,
New York, USA.
\texttt{cjiahao@gmail.com}
}
 and
Weijian Zhang
\thanks{%
  School of Mathematics,
The University of Manchester,
                Manchester, M13 9PL, England.
\texttt{weijian.zhang@manchester.ac.uk}.
}
}

\def\R{\mathbb{R}}
\def\C{\mathbb{C}}
\def\nbyn{n \times n}
\def\mbyn{m \times n}
\def\l{\lambda}
\def\norm#1{\|#1\|}
\def\normi#1{\|#1\|_1}
\def\normo#1{\|#1\|_{\infty}}
\def\Chat{\widehat{C}}
\def\e{eigenvalue}

% \DeclareMathOperator{\diag}{diag}   % Requires amsmath.
\def\diag{\mathop{\mathrm{diag}}}     % If not using amsmath.
\def\trace{\mathop{\mathrm{trace}}}   % If not using amsmath.

\def\At{\widetilde{A}}
\def\normt#1{\|#1\|_2}

% Note: this clashes with amsmath.
\def\bmatrix#1{\left[\matrix{#1}\right]}

\newenvironment{njh}{\begin{quote}\small\sf{\color{blue}$\diamondsuit$~NJH~}}{\end{quote}}
\newenvironment{wz}{\begin{quote}\small\sf{\color{green}$\diamondsuit$~WZ~}}{\end{quote}}

\begin{document}

%\subtitlenote{The full version of the author's guide is available as
 % \texttt{acmart.pdf} document}

\maketitle

\begin{abstract}
This paper studies the influence of the temporal information in an evolving graph on the node centrality scores.
In particular, we look at four important centrality measures, namely Katz centrality,
Resolvent Betweenness centrality, Closeness centrality, and PageRank.

We derive evolving graph centrality measures from the perspective of network flow and
observe that the amount of temporal distance between nodes has an important impact on the final evolving graph centrality scores, especially on the top ranked nodes.
Build on this idea, we introduce a generalised PageRank on evolving graph that is based on two ideas:
reverse network flow and block adjacency matrix.
We conduct our experiments on random evolving graphs to illustrate our idea.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

In the 1985 science-fiction film ``Back to the Future'', Marty McFly (played by Michael J. Fox) travelled back in time and accidentally changed history.
However, we cannot travel back in time or even send a message to the past.
To traverse an evolving graph, a.k.a a time-dependent graph, one has to consider order of time stamps. In particular, a walk on an evolving graph needs to be time-preserving, meaning we
can not pass a massage (through an edge) to a node at a earlier time.
Time-preversing walks and paths are very important concepts for
network flow on evolving graphs.

% time-respecting walk
We represent an evolving graph as a time-ordered sequences of graphs, similar to the work of Tang and coworkers \cite{nicosia13, tang09, tang102, tang10} and Grindrod, Higham and coworkers \cite{grindrod13,grindrod11}. The idea is to divide the system's time span into temporal slices and regards it as a sequence of static graphs $G^{[t]}$, one for each layer. Such representation arises naturally in many real-world applications. For example,
consider networks of users interacting through messaging. Each interaction between two users $A$ and $B$ are stored with a time stamp $t$. If we represent a user as a node then a user $A$ at time stamp $t$ is represented by a pair $(v,t)$ and a message sent from user $A$ to user $B$ at time stamp $t$ can be represented as $\langle (A, t) (B, t) \rangle$.

In general, a node $v$ on $G^{[t]}$ is represented by a pair $(v, t)$, where $t$ is the time stamp of the graph $G^{[t]}$.
A sequence of interactions between users form a walk. Suppose $A$ sent a message to $B$ at time stamp $t_1$
and then $B$ sent this message to $C$ at time stamp $t_2$. We can represent this activity as $\langle (A,t_1), (B,t_1), (B, t_2), (C,t_2)\rangle$, where each pair of two nodes has meanings: $\langle (A,t_1), (B,t_1) \rangle$ and $\langle (B,t_2), (C,t_2) \rangle$ represent interactions between users $A$, $B$, and $C$ and $\langle (B,t_1), (B,t_2) \rangle$ represents the fact that user $B$ holds the message at time stamp $t_1$ and $t_2$.

A time-respecting walk is a sequence of nodes $\langle (v_1, t_1), (v_2, t_2), \ldots ,(v_m, t_m)\rangle$, where $t_1 \le t_2 \le \cdots \le t_n$ are the corresponding time stamps. If $t_i = t_j$, then $\langle (v_i, t_i), (v_j, t_j) \rangle$ is an edge where both nodes are on the same temporal slice,  otherwise it is an edge linking nodes from different temporal slices.
We call edges of the first kind \emph{static edges} and of the second kind \emph{causal edges}.
Recall in static graph, a walk of length $m$ is defined as a sequence of nodes $v_1, v_2, \ldots, v_{m+1}$ such that for each $i = 1, \ldots, m$ there is an edge from $v_i$ to $v_{i+1}$. How to determine the length of a time-respecting walk?
One could define the length of a time-respecting walk as the total number of temporal slices (see shortest temporal distance in \cite{tang09}) or the number of static edges (see dynamic walk in \cite{grindrod11}). This paper studies the impact of these two different definitions. In particular, we studies a general definition of temporal distance that takes account of both temporal slices and static edges.

% Related works
Researchers have generalised many static graph centrality algorithms for evolving graphs. One approach is to use the fact that the elements of the matrix product of adjacency matrices at different time stamps can correctly count the number of temporal paths.
However, not every centrality algorithm can be generalised in this way. Another approach is to define evolving graph centrality algorithms using time-respecting paths. This approach is applicable to algorithms that are defined using shortest paths.

% no work on pagerank so far. By understanding temporal walk, we could provide a generalization of pagerank on evolving graphs.

% In Higham's following up papers, time delay is considered as alternative format of Katz centrality.
% But the impact of temporal is left unconsidered. We studied the problem on centrality from two different types of centrality: one that depends on matrix unfolding and one that depends on temporal walk.

% our contribution this paper
Borgatti \cite{borgatti05} noted that the manner of traffic flow entails a new way to think about centrality. In particularly, the importance of a node in a network can not be determined without reference to how traffic flows through the network.
Inspired by this idea, we simulate network flows in evolving graphs to measure the centrality scores.
% Grindrod et al.'s definition of dynamic walk \cite{grindrod11} only counts static edges.
% Tang et al. \cite{tang10s} measure a temporal shortest walk in time unit (or causal edges) only.
% If a node can be reached by another node in $k$ time stamps then the distance between the two nodes is $k$. Chen and Zhang \cite{chen16} considered both static edges and causal edges in their definition of temporal walk.
The aims of this paper are as follows. First, to derive important evolving graph centrality from first principle via network flow. Second, to study the impact of temporal distance between nodes on centrality.
Third, to generalise PageRank on evolving graphs.
We carry out the experiment using the Julia dynamic network analysis package
EvolvingGraphs.jl (\url{https://github.com/EtymoIO/EvolvingGraphs.jl}).
We use the research paper data gathered from Etymo (\url{https://etymo.io}), a visual search engine for data scientists.

\section{Motivation: A Closer Look at Katz Centrality on Evolving Graphs}
\label{sec:motivation}

We let $G_n = \langle G^{[1]}, G^{[2]}, \ldots, G^{[n]} \rangle$ be an evolving graph
and $A^{[1]}, A^{[2]}, \dots A^{[n]}$ be the corresponding adjacency matrices.
Then the dynamic communicability matrix is defined as
\begin{equation}
  \label{eq:q}
  Q^{[k]} = (I - \alpha A^{[1]})^{-1}(I - \alpha A^{[2]})^{-1} \cdots (I -  \alpha A^{[n]})^{-1},
\end{equation}
where parameter $\alpha < 1/ \max_k \rho(A^{[k]})$, the spectral radius of $A^{[k]}$.
The $i$th row sum of $Q^{[k]}$ is the broadcast centrality.

Suppose $A$, $B$, and $C$ are three co-workers in a large internet company. Figure \ref{fig:katz_eg}
shows how they exchange ideas during $3$ days at work. The directions of the arrows indicate the direction of the flow of ideas.
For example, $A$ shares a new idea with $B$ on day $1$ and $B$ shares a new idea with $C$ on dat $3$. As a result, $A$'s idea can influence $C$.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n10/n7, n1/n4, n4/n1, n11/n12, n1/n6}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{$2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $3$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $2$ and $3$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg}
\end{figure}

We could use EvolvingGraphs.jl to generate the above evolving graph and compute the Katz centrality.

\begin{lstlisting}
using EvolvingGraphs
using EvolvingGraphs.Centrality

g = EvolvingGraph{Node{String}, Int}()
add_bunch_of_edges!(g, [("A", "B", 1), ("A", "C", 2),
("A", "B", 2),("C", "A", 2),("B", "C", 3)])
\end{lstlisting}

The centrality values for each node are
\begin{lstlisting}
katz(g)
3-element Array{Tuple{EvolvingGraphs.Node{String},Float64},1}:
 (Node(A), 0.742301)
 (Node(B), 0.42943)
 (Node(C), 0.514373)
\end{lstlisting}

As expected $A$ is the most important node in the network. $C$ is the second important node in the network as it influenced
$A$ at time stamp $2$. If we reverse the time of communication between day $1$ and $3$, i.e., $B$ shares a new idea with $C$ on day $1$ and $A$ shares a new idea with $B$ on day $3$. Then in this case the idea $A$ had on day $3$ can not pass to $C$. This is illustrated in Figure \ref{fig:katz_eg2}.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n7/n8, n1/n4, n4/n1, n1/n6, n14/n11}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $3$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $2$ and $3$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg2}
\end{figure}

In this cast, we have
\begin{lstlisting}
g2 = EvolvingGraph{Node{String}, Int}()
add_bunch_of_edges!(g2, [("A", "B", 3), ("A", "C", 2),
("A", "B", 2),("C", "A", 2),("B", "C", 1)])
\end{lstlisting}

The centrality ratings of the three nodes are

\begin{lstlisting}
katz(g2)
3-element Array{Tuple{EvolvingGraphs.Node{String},Float64},1}:
 (Node(A), 0.687679)
 (Node(B), 0.490062)
 (Node(C), 0.535666)
\end{lstlisting}

Notice the rating of $A$ decreased as its influence is not as important as in Figure \ref{fig:katz_eg}.
% does not take account of walks in a long time. cost in time
If the communication between $A$, $B$, and $C$ are not on three consecutive days
but on day $1$, day $10$, and day $100$, as shown in Figure \ref{fig:katz_eg3},
we would expect the centrality scores to change. However, according to \eqref{eq:q}, scores are unchanged.
We notice that the generalised Katz centrality on evolving graphs only takes account of the edges at each time stamp but not the communication time between edges at different time stamps.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n7/n8, n1/n4, n4/n1, n1/n6, n14/n11}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $10$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $100$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $10$ and $100$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg3}
\end{figure}

\section{Time-Preserving Walks and Paths}
\label{sec:time-pres-paths}

% what is walk and what is path?
A walk of length $l$ is a sequence of nodes $v_1, v_2, \ldots, v_l, v_{l+1}$ such that for
each $i = 1, 2, \ldots, l$, there is an edge from $v_i$ to $v_{i+1}$. A path of length $l$ is a walk of length $l$
such that all the nodes are different.
% relate to centrality
The definition of centrality depends on the manner in which traffic flows through a network \cite{borgatti05}.
Time-preserving walks induce temporal Katz centrality and temporal closeness centrality while time-preserving paths induce betweenness centrality and closeness centrality.

We now describe evolving graph network flow in more details. We follow the definition of
\emph{evolving graph} in \cite{chen16}.

\begin{definition}
  An \textbf{evolving graph} $G_n$ is a sequence of (static) graphs
$G_n = \langle G^{[1]}, G^{[2]},  \ldots ,G^{[n]} \rangle$ with associated time stamps
$t_1, t_2, \ldots, t_n$ respectively. Each $G^{[t]} = (V^{[t]}, E^{[t]})$ represents a (static) graph labelled by a time t.
\end{definition}

Note the node sets $V^{[t]}$ can change over time, i.e., nodes may appear or disappear at a particular time stamp.
For example, in Figure \ref{fig:eg_shortest_walk}, at time stamp $1998$, $V^{[1998]} = \langle 1, 2 \rangle$ and $E^{[1998]} = \langle (1,2) \rangle$. Each graph $G^{[t]}$ can be represented by its adjacency matrix $A^{[t]}$.
We can represent $G_n$ by a list of adjacency matrices $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{[n]} \rangle$.

If we only measure the temporal difference in a temporal walk, we have a new definition of Katz centrality.
In the matrix form, the nonzero entries of $A^{[t_i]} A^{[t_{i+1}]}$ counts all the temporal walks of length $1$, and
$A^{[t_i]}\cdots A^{[t_j]}$, where $i < j$ counts all the temporal walks of length $j -i$.
For example, $A^{[1]}A^{[4]}$, $A^{[1]}A^{[1]}A^{[4]}$ and $A^{[1]}A^{[2]}A^{[3]}A^{[4]}$ are all walks of length 3.
In the original Katz centrality, the `attenuation' factor $\alpha$ is imposed on the
number of paths, i.e., longer paths has smaller effect to the overall rating.
Hence the final result is the summation of all products of the form
\[
\alpha^k A^{[i]} \cdots A^{[i+k]}
\]
Note this can not be written in \eqref{eq:q} as discussed in Section \ref{sec:temp-katz-centr}.

% Another definition of temporal walk was introduced by \cite{chen16}. Here
% both static edges and causal edges are considered. We can represent an evolving graph by a block adjacency matrix, as shown in Section \ref{sec:centr-block-adjac}. Centrality on this block matrix gives the importance of each node at all time stamps. We can average these rating at different time stamps to
% get the final rating.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
  \node (w1) at (-6, 9) {$w_1$};
  \node (w2) at (3, 13) {$w_2$};
  \node (w3) at (-2, 8) {$w_3$};
  \node (w4) at (2, 4) {$w_4$};
  \node (w5) at (8.5, 7) {$w_5$};s
  \foreach \from/\to in {n10/n7, n1/n4, n11/n12}
   \draw[every edge,->] (\from) -- (\to);
 \draw[dotted,->](n4) -- (n12);
\draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1998$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $1999$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $2018$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1998$, $1999$ and $2018$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows and directed edges between graphs are shown as dotted arrows, where $w_1, w_2 \ldots w_5$ are edge weights.}
\label{fig:eg_shortest_walk}
\end{figure}

\begin{definition}
  A \textbf{temporal node} is a pair (v, t), where $v \in V^{[t]}$ is a node at a time $t$.
\end{definition}

%At each time stamp, we only count temporal nodes that connects by an edge.

\begin{definition}
  A temporal node $(v, t)$ is an \textbf{active node} if there exists at least one edge $e \in E^{[t]}$ that connects $v \in V^{[t]}$ to another node $w \in V^{[t]}, w\ne v$. Otherwise it is called an inactive node.
\end{definition}

In Figure \ref{fig:eg_shortest_walk}, the filled green circles are active nodes. For the adjacency matrix representation, a node $i$ is active at time stamp $t$, then
the $i$th row or the $i$th column of $A^{[t]}$ has at least one none-zero entry.

 \begin{definition}
 A \textbf{static edge} $\langle (v_i, t), (v_j, t)\rangle$ is a pair of elements of $V^{[t]}$, i.e., $v_i \in V^{[t]}$ and
 $v_j \in V^{[t]}$.
  A \textbf{causal edge} $\langle (v, t_i), (v, t_j)\rangle$ is a pair of the same node at different time stamps.
 \end{definition}

 For example, in Figure \ref{fig:eg_shortest_walk}
 $\langle (1, 1998), (2, 1998) \rangle$ is a static edge at time stamp $1998$ and $\langle (3, 1999), (3, 2018) \rangle$ is a causal edge between time stamp $1999$ and time stamp $2018$.
 The difference between static edges and causal edges were first explicitly considered in \cite{chen16}.

 We can denote a \emph{weighted static edge} as $\langle (v_i, t), (v_j, t), w_s \rangle$ and a \emph{weighted causal edge} as $\langle (v, t_i), (v, t_j), w_t \rangle$.
 Here $w_s$ represents the spatial distance between the two nodes $v_i$ and $v_j$; $w_t$ represents the
 temporal distance of $v$ at different time stamps.
 In Figure \ref{fig:eg_shortest_walk}, the black arrows represent static edges and the dotted arrows represent causal edges.
 Suppose we assign all the static edge weight to $1$ and
 let the causal edge weight be the number of time stamps between the pair of nodes. Then in Figure \ref{fig:eg_shortest_walk}
 the edge weight between $(1, 1998)$ and $(2, 1998)$ is $1$ and the edge weight between $(2, 1998)$ and $(2, 2018)$ is $10$.
In general, a time-preserving walk can be defined as follows.

\begin{definition}
A \textbf{time-preserving walk}  of length $m$ on an evolving graph $G_n$ is a time-ordered sequence of active nodes, $\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$, where $t_1 \le t_2 \le \cdots \le t_m$ and
$v_i = v_j$ if and only if $t_i \ne t_j$.
\end{definition}

It is natural to define the length of a walk on a static graph as the number of nodes travelled through the walk. For the weighted graph case, the length of a walk is the total weight of all the edges in the walk. We observe that static edges link nodes in space while causal edges link nodes in time. Hence we would like to distinguish temporal distance from spatial distance in an evolving graph: we let the length of a static edge be one and the length of a causal edge be the temporal distance between the two nodes. In the simplest case, the time-respecting walk
$\langle (1, 1998) ,(2, 1998) , (2, 2018), (3, 2018)\rangle$ in Figure \ref{fig:eg_shortest_walk} has length $12$ in total because two static edges $\langle (1, 1998) (2, 1998) \rangle$ and $\langle (2, 2018), (3, 2018) \rangle$ each have length one, and $\langle (2, 1998) (2, 2018) \rangle$
traverse $10$ time stamps and thuse has length $10$.

A time-preserving walk (and path) can traverse both causal edges and static edges.
The following Julia function measures the temporal distance between two active nodes in an evolving graph.

\begin{lstlisting}
function temporal_distance(v1::TimeNode,
                           v2::TimeNode, beta::Real = 1.)
  beta* abs(node_timestamp(v1) - node_timestamp(v2))
end
\end{lstlisting}
which means the temporal distance is the difference between their time stamps scaled a hyper-parameter
\texttt{beta}.

\section{Centrality Measures}
\label{sec:topol-temp-flow}

% what is centrality?
Centrality measures the importance of nodes within a graph.
% current evolving graph centrality approach
To generalize centrality measures for evolving graphs there are two common approaches.
First, we could exploit the fact that matrix product of adjacency matrices at different time stamps counts the number of dynamic walks in an evolving graph. Using this idea, we can generalize walk-based centrality measures such as
the Katz Centrality \cite{estrada09} and Communicability Betweenness \cite{alsayed15}.
Second, we could generalize the definition of shortest paths as shortest temporal paths and replace shortest paths in centrality measures with shortest temporal paths where possible. This way we can define temporal betweenness centrality and temporal closeness centrality \cite{nicosia13}.

% our approach
We observe both approaches ignore the communication time between edges at different time stamps, i.e., the time gap between different (static) graph slices. We take account of this communication time by deriving centrality measures from the original theses of the ideas and show communication time has important impact on the final node ratings (and rankings).
We note that our goal is not about providing efficient centrality algorithms which is out of the scope of this paper.

\subsection{Temporal Katz Centrality}
\label{sec:temp-katz-centr}

% katz centrality from first principle
Katz Centrality accounts the influence of nearest-neighbours to a give node and the influence of other nodes
separated at a certain distance from it. The $k$th power of adjacency matrix $A$, i.e., $A^k$, of a graph $G$ accounts for walk of length $k$. The expansion
\begin{equation}
  \label{eq:a}
  A^0 + \alpha A + \alpha^2 A^2 + \cdots,
\end{equation}
converges to $(I - \alpha A)^{-1}$ when $\alpha < 1/\rho(A)$, the spectral radius of $A$. The $(i,j)$ entry of \eqref{eq:a}
accounts all walks from $i$ and $j$ with the influence of walks of length $k$ scaled by a factor of $\alpha^k$.
The Katz centrality of node $i$ is the $i$th row sum of $(I - \alpha A)^{-1}$, which is the sum of influence of all the nodes that node $i$ can reach via a walk.
In other words, we start with node $i$ and each out-neighbour of node $i$ has influence $\alpha$ on node $i$ and the out-neighbour of out-neighbour of node $i$ has influence $\alpha^2$ on node $i$, and so on.

For evolving graphs, we replace out-neighbours with forward neighbours which preserve the direction of time.
Then the Katz score of a node $i$ at time stamp $t$, denoted by $(i,t)$, is the sum of influence of all the active nodes that $(i,t)$ can reach via a time-preserving walk.
Unlike \cite{grindrod11}, we also consider the temporal distance between two nodes. For example, the temporal distance between $(i, 2001)$ and $(i, 2018)$ is $17$.
% julia code implementation
With the definition of \texttt{temporal\_distance} introduced in Section \ref{sec:time-pres-paths}.
We define the temporal Katz score of node $(i, t)$ as

\begin{lstlisting}[caption={Temporal Katz Centrality of Single Node},label={lst:katz}]
function temporal_katz(g::AbstractEvolvingGraph,
  start::TimeNode; alpha = 0.2, max_level = 10)
    score = 0.
    v = start
    fronter = [v]
    level = 0
    while level < max_level
        next = []
        for u in fronter
            for v in forward_neighbors(g, u)
                push!(next, v)
                if node_key(v) != node_key(u)
                  td = temporal_distance(start, u)
                  d = td + level
                  score += alpha^d
                end
            end
        end
        fronter = next
        level += 1
    end
    return score / num_edges(g)
end
\end{lstlisting}

% explain the code
The algorithm \texttt{temporal\_katz} performs a breadth first search (BFS) on evolving graph $g$ that
starts at \texttt{TimeNode} $v$, which represents a node at a specific time stamp \footnote{See also \url{https://etymoio.github.io/EvolvingGraphs.jl/latest/base.html#EvolvingGraphs.TimeNode}}.
For each \texttt{TimeNode} $v$, we accumulate the influence of linked nodes according to their spatial and temporal
distance from $v$. The spatial distance is recorded in variable \texttt{level} and the temporal distance is calculated by
\texttt{temporal\_distance}. The algorithm stops searching when \texttt{level} is larger or equal to \texttt{max\_level}.
We set \texttt{alpha} to be $0.2$ and \texttt{max\_level} to be $10$. Then
% example on 3 timestamp evolving graph
for Figure \ref{fig:katz_eg2}, we have
\begin{lstlisting}
  TimeNode(A, 2) => 0.466667
  TimeNode(A, 3) => 0.2
  TimeNode(B, 2) => 0.0
  TimeNode(B, 3) => 0.0
  TimeNode(B, 1) => 0.202347
  TimeNode(C, 1) => 0.0117333
  TimeNode(C, 2) => 0.293333
\end{lstlisting}
The total scores at all time stamps are
\begin{lstlisting}
  "A" => 0.666667
  "B" => 0.202347
  "C" => 0.305067
\end{lstlisting}
Notice that $A$ is the most important node in the evolving graph and $C$ is the second most important node.
Thus the ranking agrees with the Katz centrality rankings in Section \ref{sec:motivation}.

For Figure \ref{fig:katz_eg3}, we have
\begin{lstlisting}
  TimeNode(A, 10)  => 0.458333
  TimeNode(A, 100) => 0.2
  TimeNode(B, 1)   => 0.2
  TimeNode(B, 10)  => 0.0
  TimeNode(B, 100) => 0.0
  TimeNode(C, 1)   => 2.98666e-8
  TimeNode(C, 10)  => 0.291667
\end{lstlisting}
The sum of node scores at each time stamp is
\begin{lstlisting}
  "A" => 0.658333
  "B" => 0.2
  "C" => 0.291667
\end{lstlisting}
We see the scores of all three nodes are decreased when it takes longer to communicate between different time stamps. In this case, the overall ranking is not affected but in general the communication time between time stamps can change the overall ranking.

\subsubsection{Temporal Resolvent Betweenness Centrality}

Betweenness centrality measures the importance of a node as the ability to facilitate the communication among other nodes in the network. Commonly it is defined based on the shortest paths. However, key messages do not necessarily follow the shortest paths and thus it makes senses to consider walks instead of shortest paths.
Recall the $(i,j)$ entry of the resolvent matrix $(I - \alpha A)^{-1}$ provide information about the communication from $i$ to $j$. We could therefore define the resolvent betweenness for node $v$ as
\begin{equation}
  \sum_{i}\sum_j \frac{(I-\alpha A)^{-1}_{i,j} - (I - \alpha (A - E_v))^{-1}_{i,j}}{(I - \alpha A)^{-1}_{i,j}}, \quad i \ne j, i \ne v, j \ne v.
\end{equation}
where $E_v$ has non-zeros only in row and column $v$, and in the $v$th row and column has $1$ where $A$ has $1$.
The matrix $A - E_i$ represents a graph with all edges involving the node $v$ removed.
For $(I-\alpha A)^{-1}_{i,j}$ we could simply modify Listing \ref{lst:katz} so that it only accumulate scores when the end node is $j$. For the $(I - \alpha (A-E_v))^{-1}_{i,j}$ part, we can apply the above the same algorithm on a graph with node $v$ and related edges removed.
Therefore the general temporal resolvent betweenness centrality can be defined as
\begin{lstlisting}
function temporal_resolvent_betweenness(g1, g2,
  v::TimeNode; alpha = 0.2, k = 10)
    r = 0.
    ns = active_nodes(g1)
    for i in ns
        for j in ns
            if i != j && i != v && j != v
                score1 = temporal_katz(g1, i, j,
                    alpha = alpha, k = k)
                score2 = temporal_katz(g2, i, j,
                    alpha = alpha, k = k)
                r += score1 == 0.0 ? 0 :
                    (score1 - score2)/score1
            end
        end
    end
    return r
end
\end{lstlisting}
Here \texttt{temporal\_katz} only takes account of walks from node $i$ to $j$. Evolving graph
\texttt{g1} is the original evolving graph and evolving graph \texttt{g2} is \texttt{g1} with all the edges
related to \texttt{v} removed.

For example, by removing node $(A, 2)$, i.e., node $A$ at time stamp $2$, from Figure \ref{fig:katz_eg2} we find
the temporal resolvent betweenness score of node $A$ is $5.0$. While the temporal resolvent betweenness score of node $(B,3)$ is $4.0$. This shows that $(A,2)$ is more important than $(B,3)$.
We note that temporal communicability betweenness centrality can be analysed similarly.

\subsection{Temporal Closeness Centrality}
\label{sec:temp-betw-centr}

% closeness centrality

Let $d_{ij}$ be the length of the shortest path from $i$ and $j$ in a static graph. Then
the closeness centrality of a node $i$ is defined
$$
C_i^{closeness} = \frac{N-1}{\sum_j d_{ij}},
$$
where $N-1$ is the total number of reachable nodes.
For evolving graphs, we need to consider the shortest temporal path from $(i,t_i)$ to $(j,t_j)$, for any time stamp $t_i \leq t_j$.
Depends on how define the length of shortest temporal path, the closeness centrality can counts static edges, causal edges, or both.
The key difference between Katz centrality and closeness centrality is the fact here we only consider shortest temporal paths, not just every temporal walks.
We could apply BFS to find the shortest temporal path from $(v,t)$ to any other nodes in the evolving graphs. The algorithm is shown as follows.

\begin{lstlisting}
function temporal_closeness(g::AbstractEvolvingGraph,
  start::TimeNode)
  v = start
  level = Dict(v => 0)
  i = 1
  fronter = [v]
  while length(fronter) > 0
      next = []
      for u in fronter
          for v in forward_neighbors(g, u)
              if !(v in keys(level))
                  if node_key(u) != node_key(v)
                    td = temporal_distance(start, u)
                    level[v] = i + td
                  else
                    level[v] = i - 1
                  end
                  push!(next, v)
              end
          end
      end
      fronter = next
      i += 1
  end

  total_scores = sum(values(level))
  return total_scores > 0. ?
      (length(level) - 1)/total_scores : 0.
end
\end{lstlisting}
The algorithm starts at node \texttt{start} and explores the forward neighbours and store all the
reached nodes so far in \texttt{level}. If \texttt{node\_key(u) == node\_key(v)}, we traverse the same node at different time stamps. In this case, we consider that we've already reached it at the previous time stamp.

For Figure \ref{fig:katz_eg2} the scores are
\begin{lstlisting}
TimeNode(A, 2) => 0.666667
TimeNode(A, 3) => 1.0
TimeNode(B, 1) => 0.346154
TimeNode(B, 2) => 0.0
TimeNode(B, 3) => 0.0
TimeNode(C, 1) => 0.315789
TimeNode(C, 2) => 0.5
\end{lstlisting}
and the sum of scores at different time stamps gives
\begin{lstlisting}
"A" => 1.66667
"B" => 0.346154
"C" => 0.815789
\end{lstlisting}
Note that the ranking is the same as the temporal Katz centrality case.
For Figure \ref{fig:katz_eg3}, the scores are
\begin{lstlisting}
TimeNode(A, 10)  => 0.0707071
TimeNode(A, 100) => 1.0
TimeNode(B, 1)   => 0.0597015
TimeNode(B, 10)  => 0.0
TimeNode(B, 100) => 0.0
TimeNode(C, 1)   => 0.0390625
TimeNode(C, 10)  => 0.0412371
\end{lstlisting}
and the sums are
\begin{lstlisting}
"A" => 1.07071
"B" => 0.0597015
"C" => 0.0802996
\end{lstlisting}
Similar to the temporal Katz centrality case, we see a decrease of centrality scores
for all the nodes in the evolving graph.
We note that the analysis for temporal betweenness centrality is similar to above.

\subsection{Temporal PageRank}

We see from our analysis for temporal Katz centrality and temporal betweenness centrality
that the temporal distance between nodes in an evolving graph can impact the final centrality scores.
We would like to design a temporal PageRank algorithm that takes account of between
the temporal distance and the spatial distance. The generalisation is based on two ideas:
reversed temporal walks and block adjacency matrix.

\subsubsection{Reversed Temporal Walks}

We can not travel back in time. But reversing the direction of temporal flows can help
identify the ``hub'' nodes or the source of the flows. Recall in the HITS algorithm \cite{kleinberg99},
the hub score describes the quality of a web page as a link collection of important related pages.
In reverse PageRank we compute PageRank on the graph with reversed direction, i.e., reverse the direction of each edge $(i,j)$ to $(j,i)$. Fogaras \cite{fogaras03, gleich15} shows that Reversed Page Rank scores express hub quality.

Recall a time-preserving walk is a time-ordered sequence of active nodes. We define a \emph{reversed temporal walk} to be a reversed time-ordered sequence of active nodes
$\langle (v_1, t_1), (v_2,t_2), \ldots, (v_m, t_m)\rangle$, where
$t_1 \ge t_2 \ge \cdots \ge t_m$.

\subsubsection{Block Adjacency Matrices}
\label{sec:centr-block-adjac}

Aggregating edges at each time stamp to form
a simple static graph can provide misleading information about the structure of the network.
In \cite{chen16}, we derive a block adjacency matrix representation of an evolving graph and prove it is the correct representation of an evolving graph.
Let $E'$ is the set of causal edges and $\hat E$ is the set of static edges then
an evolving graph can be represented as
$$
\bm M_n =
\begin{pmatrix}
A^{[t_1]} & M^{[t_1, t_2]} & \ldots & M^{[t_1, t_n]} \\
0         & A^{[t_2]} & \ldots & M^{[t_2, t_n]} \\
          & \ldots    &        &     \\
0         & 0         & \ldots & A^{[t_n]}
\end{pmatrix},
$$
where $M^{[t_i, t_j]}$ is the matrix whose rows are labeled by $V^{[t_i]}$ and columns are labeled by $V^{[t_j]}$, and whose entries are
$$
  M_{uv}^{[t_i, t_j]} =
  \begin{cases}
    1 & \mbox{if} (u, v) \in E' \\
    0 & \mbox{otherwise}.
  \end{cases}
$$
The adjacency matrix blocks $A^{[t]}$ encode the static edge set $\hat E$, whereas the off-diagonal blocks $M^{[t_i, t_j]}$ encode the causal edge set $E'$. To take account of the temporal distance between nodes, we could let
$$
  M_{uv}^{[t_i, t_j]} =
  \begin{cases}
    \frac{1}{|t_j - t_i|} & \mbox{if} (u, v) \in E' \\
    0 & \mbox{otherwise}.
  \end{cases}
$$

For Figure \ref{fig:eg_shortest_walk}, we have

\begin{align*}
V & = \{(1,\!1998), (2,\!1998), (1,\!1999), (3,\!1999), (2,\!2018), (3,\!2018)\},\\
\tilde E &= \{((1,\!1998), (2,\!1998)), ((1,\!1999), (3,\!1999)), ((2,\!2018), (3,\!2018))\},\\
E'  &= \{((1,\!1998), (1,\!1999)), ((2,\!1999), (2,\!2018)), ((3,\!1999), (3,\!2018))\}.
\end{align*}

and corresponding block adjacency matrix is

\[
\bm M_3 = \begin{pmatrix}
0 & 1 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1/9 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 1/9 \\
0 & 0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 0 & 0 & 0
\end{pmatrix}
\]

The static graph representation of Figure \ref{fig:eg_shortest_walk} is shown in Figure \ref{fig:static}.

\begin{figure}[h]
 \begin{center}
   \begin{tikzpicture}[scale=.36, line width =1.4pt]
  \node[circle,fill=green1, minimum size=0.1cm, inner sep = 0pt] (n7) at (-5,7)
{\scriptsize $(2,t_1)$};
  \node[circle,draw=red1, minimum size=0.2cm, inner sep = 0pt] (n8) at (-10,5)
{\scriptsize $(3,t_1)$};
  \node[circle,fill=green1, minimum size=0.1cm, inner sep = 0pt] (n10) at (-8,10) {\scriptsize $(1,t_1)$};

  \node[circle,draw=red1, minimum size=0.2cm, inner sep = 0pt] (n6) at (3,7)
{\scriptsize $(2,t_2)$};
  \node[circle,fill=green1, minimum size=0.2cm,  inner sep = 0pt] (n4) at (-2,5)
{\scriptsize $(3,t_2)$};
  \node[circle,fill=green1, minimum size=0.2cm,  inner sep = 0pt] (n1) at (0,10)
{\scriptsize $(1,t_2)$};

   \node[circle,fill=green1, minimum size=0.2cm,  inner sep = 0pt] (n11) at (11,7)
{\scriptsize $(2,t_3)$};
  \node[circle,fill=green1, minimum size=0.2cm,  inner sep = 0pt] (n12) at (6,5)
{\scriptsize $(3,t_3)$};
  \node[circle,draw=red1, minimum size=0.2cm,  inner sep = 0pt] (n14) at (8,10)
{\scriptsize $(1,t_3)$};

  \foreach \from/\to in {n10/n7, n1/n4, n11/n12}
   \draw[every edge,->] (\from) -- (\to);
     \draw[dotted,->](n7) to[out=80, in=40] (n11);
  \draw[dotted,->](n10) to[out=10, in=180] (n1);
   \draw[dotted,->](n4) to[out=-30,in=-150] (n12);
    \end{tikzpicture}
\end{center}
\caption{A static graph corresponding to the evolving graph example of
Figure~\ref{fig:eg_shortest_walk}. The green nodes are active nodes while the
red nodes are inactive nodes.
The black lines are edges in the static edge set $\tilde E$ and are encoded
algebraically in the diagonal blocks $A^{[t]}$ of the adjacency matrix $\bm M_3$.
The dotted lines are edges in the causal edge set $E'$ and are encoded algebraically in
the off-diagonal blocks $M^{[t_i, t_j]}$.
The graph containing all the edges and temporal nodes has adjacency matrix $\bm M_3$.}
\label{fig:static}
\end{figure}

Notice that the lower triangular blocks of matrix $\bm M_n$ are zeros.
This is because all the causal edges are time preserving. We consider the transpose of matrix $\bm M_n$, which reverse the direction of static edges and causal edges.
Let $b \in [0,1]$ be a balance scalar. Then the block matrix
$$
b \bm M_n + (1-b) \bm M_n^T
$$
takes account of both the temporal walks and reversed temporal walks.
We now derive the block Google matrix as

\begin{lstlisting}
function block_google_matrix(g; alpha = 0.85, balance = 0.75)
  M = full(block_adjacency_matrix(g))
  M = balance * M + (1-balance) * M'
  N, N = size(M)
  p = ones(N)/N
  dangling_weights = p
  dangling_nodes = find(x->x==0, sum(M,2))
  for node in dangling_nodes
      M[node,:] = dangling_weights
  end
  M ./= sum(M,2)
  return alpha * M .+ (1- alpha) * p
end
\end{lstlisting}

The function \texttt{block\_adjacency\_matrix} convert an evolving graph $g$ to a block adjacency matrix $\bm M$. Every row sum of the block Google matrix is equal to one. Each element of the dominant eigenvector of the block Google matrix
represents the PageRank rating of a node.

For Figure \ref{fig:katz_eg2}, we set $alpha$ be $0.85$ and $balance$ be $1.0$.
Then the PageRank scores are
\begin{lstlisting}
 TimeNode(A, 2) => 0.289123
 TimeNode(A, 3) => 0.128054
 TimeNode(B, 1) => 0.182477
 TimeNode(B, 2) => 0.354335
 TimeNode(B, 3) => 0.480941
 TimeNode(C, 1) => 0.128054
 TimeNode(C, 2) => 0.250931
\end{lstlisting}
and the sum of node scores at all three time stamps are
\begin{lstlisting}
 "A" => 0.417177
 "B" => 1.01775
 "C" => 0.378986
\end{lstlisting}
We see node $B$ is the most important node in an evolving graph.
On the other extreme, if we set $balance$ to be $0.0$, we get the overall scores
\begin{lstlisting}
 "A" => 0.678502
 "B" => 0.647353
 "C" => 0.527927
\end{lstlisting}
Thus $A$ is the most important node in this case.

One potential problem with temporal PageRank is that the computational cost is expensive for large evolving graphs with many time stamps. The dimension of the block matrix
is $|V| \times |T|$, where $|V|$ represents the number of nodes and $|T|$ represents the number of time stamps.
\section{Experiments}
\label{sec:experiments}

We conduct the following experiments using \texttt{IntAdjacencyList}, which represents an evolving graph $g$ with integer nodes and integer time stamps as adjacency lists. Each active node stores its forward neighbours as a list. We count both static edges and causal edges.
The following experiments study the influence of temporal distance on centrality scores. All experiments are conducted on a MacOS system with $8$ GB of RAM running at $2.3$ GHz clock speed.

\subsection{Random Evolving Graphs}
\label{sec:random}

We generate a random evolving graph $g$ with $500$ nodes and about $499183$ static edges and $10$ time stamps.
Here are the top $10$ nodes we find using the Katz Centrality

\begin{lstlisting}
(459, 0.126296)
(147, 0.116541)
(326, 0.113667)
(183, 0.108705)
(469, 0.0936847)
(296, 0.0925748)
(424, 0.0921004)
(377, 0.091973)
(313, 0.0910352)
(127, 0.089123)
\end{lstlisting}

Since we are usually interested in the ranking of top score nodes,
we apply temporal Katz centrality on these above $10$ nodes.
We set $alpha = 0.5$  and $max\_level = 3$. We report the change of ranking and rating in Table \ref{table:ranking}

  \begin{table}[h]
    \begin{center}
  \begin{tabular}{ c | c | c | c }
    Ranking at $t_1$ & Rating at $t_1$ & Overall ranking & Overall rating \\
     469 & $0.7193$ & 183 & $5.975$ \\
     459 & $0.7180$ & 469 & $5.932$ \\
     424 & $0.7108$ & 326 & $5.804$ \\
     296 & $0.7042$ & 313 & $5.797$ \\
     326 & $0.6853$ & 459 & $5.788$ \\
     183 & $0.6647$ & 147 & $5.782$ \\
     377 & $0.6622$ & 377 & $5.691$ \\
     127 & $0.6316$ & 296 & $5.681$ \\
     147 & $0.6303$ & 424 & $5.636$ \\
     313 & $0.5715$ & 127 & $5.613$ \\
  \end{tabular}
  \end{center}
  \caption{Temporal Katz centrality ranking and rating of evolving graph $g$.
We set the weight of all causal edges to zero.}
  \label{table:ranking}
  \end{table}

  \begin{table}[h]
    \begin{center}
  \begin{tabular}{ c | c | c | c }
    Ranking at $t_1$ & Rating at $t_1$ & Overall ranking & Overall rating \\
     \textbf{459} & $0.6290$ & 183 & $5.576$ \\
     \textbf{469} & $0.6281$ & 469 & $5.533$ \\
     424 & $0.6236$ & 326 & $5.414$ \\
     296 & $0.6167$ & 313 & $5.409$ \\
     326 & $0.5983$ & 459 & $5.401$ \\
     183 & $0.5778$ & 147 & $5.396$ \\
     377 & $0.5775$ & 377 & $5.311$ \\
     127 & $0.5492$ & 296 & $5.299$ \\
     147 & $0.5471$ & 424 & $5.258$ \\
     313 & $0.4916$ & 127 & $5.237$ \\
  \end{tabular}
  \end{center}
  \caption{Temporal Katz centrality ranking and rating of evolving graph $g$.
  We set the weight of a causal edge between two active nodes to be their temporal difference.
  Compare with Table \ref{table:ranking}, we highlight the ranking difference using bold font.}
  \label{table:ranking2}
  \end{table}

We observe the rating of all nodes are smaller when we take account of the temporal information. The only difference between the rankings of two table
is the top two nodes $459$ and $469$. The overall ranking stays unchanged.
This shows that temporal Katz centrality is stable for changing temporal information.

On the same random evolving graph $g$, we compute temporal closeness centrality.

\begin{table}[h]
  \begin{center}
\begin{tabular}{ c | c | c | c }
  Ranking at $t_1$ & Rating at $t_1$ & Overall ranking & Overall rating \\
   296 & $0.4938$ & 469 & $5.040$ \\
   469 & $0.4933$ & 183 & $5.030$ \\
   459 & $0.4915$ & 313 & $5.016$ \\
   424 & $0.4888$ & 326 & $5.009$ \\
   326 & $0.4879$ & 147 & $5.008$ \\
   377 & $0.4872$ & 459 & $5.003$ \\
   183 & $0.4858$ & 296 & $4.002$ \\
   127 & $0.4831$ & 377 & $4.996$ \\
   147 & $0.4790$ & 127 & $4.989$ \\
   313 & $0.4773$ & 424 & $4.986$ \\
\end{tabular}
\end{center}
\caption{Temporal closeness centrality ranking and rating of evolving graph $g$.
We set the weight of all causal edges to zero.}
\label{table:ranking3}
\end{table}

\begin{table}[h]
  \begin{center}
\begin{tabular}{ c | c | c | c }
  Ranking at $t_1$ & Rating at $t_1$ & Overall ranking & Overall rating \\
   296 & $0.2540$ & 469 & $3.562$ \\
   469 & $0.2487$ & \textbf{147} & $3.557$ \\
   \textbf{377} & $0.2478$ & \textbf{183} & $3.555$ \\
   \textbf{459} & $0.2475$ & \textbf{313} & $3.548$ \\
   \textbf{424} & $0.2471$ & \textbf{377} & $3.545$ \\
   \textbf{326} & $0.2470$ & \textbf{296} & $3.541$ \\
   \textbf{127} & $0.2445$ & \textbf{459} & $3.540$ \\
   \textbf{183} & $0.2422$ & \textbf{127} & $3.538$ \\
   \textbf{313} & $0.2389$ & \textbf{424} & $3.535$ \\
   \textbf{147} & $0.2373$ & \textbf{326} & $3.533$ \\
\end{tabular}
\end{center}
\caption{Temporal closeness centrality ranking and rating of evolving graph $g$.
We set the weight of a causal edge between two active nodes to be their temporal difference. Compare with Table \ref{table:ranking3}, we highlight the ranking difference using bold font.}
\label{table:ranking4}
\end{table}

For temporal closeness centrality, we notice that most of rankings are changed when we take account of temporal information.

% \subsection{JMLR Coauthor Network}
% \label{sec:jmlr-coauth-netw}
%
% We construct the evolving coauthor network from Etymo (\url{https://etymo.io}).
% We collected $1566$ publications published by Journal of Machine Learning Research (JMLR) from $2001$ to $2017$ with authors involved. We regard each year
% as a time stamp and there are $17$ time stamps in total. At each time stamp, we
% create a undirected edge (or two directed edges) between two authors if they have coauthored at least one paper.
%
% Some experiments as above.
%
% Katz score | temporal katz at with time influence 0.1 | temporal katz with time influence 1.0
%
%
% Closeness with time influence 0.1 | closeness with time influence 1.0
%
%
% PageRank with time influence 0.1 | PageRank with time influence 1.0
%
% also show trends of an author.

\section{Conclusion}
\label{sec:conclusion}

We describe in details the generalizartion of centrality algorithms for evolving graphs and find
We introduce the generalization of PageRank centrality.

\bibliographystyle{plain}
\bibliography{sigproc}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: